---
title: "ICH Q2(R1) Interference/Specificity Validation Report"
subtitle: "Pharmaceutical Analytical Method Validation"
author: "Abbott Manufacturing - Analytical Development"
date: today
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
execute:
  warning: false
  message: false
params:
  analyte: "Glucose"
  interferent: "Hemolysis"
  pool_level: "All"
  spec_limit: 2.0
  equivalence_margin: 2.0
  study_id: "INT-2026-001"
  protocol_number: "VAL-ANA-2026-042"
---

```{r setup, include=FALSE}
library(tidyverse)
library(gtsummary)
library(gt)
library(broom)

# Parameters
ANALYTE <- params$analyte
INTERFERENT <- params$interferent
POOL_LEVEL <- params$pool_level
SPEC_LIMIT <- params$spec_limit
EQUIV_MARGIN <- params$equivalence_margin
STUDY_ID <- params$study_id
PROTOCOL <- params$protocol_number

# Load data
studies <- read_csv("../data/interference_studies.csv", show_col_types = FALSE) |>
  mutate(study_date = as.Date(study_date))

limits <- read_csv("../data/decision_limits.csv", show_col_types = FALSE)

# Filter data
filtered <- studies |>
  filter(analyte == ANALYTE, interferent == INTERFERENT)

if (POOL_LEVEL != "All") {
  filtered <- filtered |> filter(pool_level == POOL_LEVEL)
}

# Get regulatory limit
limit_row <- limits |> filter(analyte == ANALYTE)
reg_limit <- if (nrow(limit_row) > 0 && !is.na(limit_row$percent_limit[1])) {
  limit_row$percent_limit[1]
} else {
  SPEC_LIMIT
}
```

**DISCLAIMER:** This project contains synthetic data and analysis created for demonstration purposes only.

# Document Control

| Field | Value |
|-------|-------|
| **Study ID** | `r STUDY_ID` |
| **Protocol Number** | `r PROTOCOL` |
| **Report Date** | `r format(Sys.Date(), "%d %B %Y")` |
| **Analyte (API)** | `r ANALYTE` |
| **Interferent/Excipient** | `r INTERFERENT` |
| **Acceptance Criterion** | ±`r SPEC_LIMIT`% |
| **Regulatory Reference** | ICH Q2(R1) Section 2.2 |

# Executive Summary

```{r executive-summary, echo=FALSE}
# Calculate statistics
stats_df <- filtered |>
  group_by(interferent_concentration) |>
  summarise(
    n = n(),
    mean_measured = mean(measured_value),
    sd_measured = sd(measured_value),
    mean_bias_pct = mean(bias_percent),
    sd_bias_pct = sd(bias_percent),
    baseline = first(baseline_value),
    .groups = "drop"
  ) |>
  mutate(
    cv_percent = (sd_measured / mean_measured) * 100,
    assessment = case_when(
      interferent_concentration == 0 ~ "Baseline",
      abs(mean_bias_pct) > SPEC_LIMIT ~ "OOS",
      abs(mean_bias_pct) > SPEC_LIMIT * 0.5 ~ "Borderline",
      TRUE ~ "Pass"
    )
  )

oos_rows <- stats_df |> filter(assessment == "OOS")
borderline_rows <- stats_df |> filter(assessment == "Borderline")
```

```{r summary-text, echo=FALSE, results='asis'}
if (nrow(oos_rows) > 0) {
  threshold <- min(oos_rows$interferent_concentration)
  max_bias <- max(abs(oos_rows$mean_bias_pct))
  direction <- ifelse(oos_rows$mean_bias_pct[1] > 0, "positive", "negative")

  cat("## RESULT: OUT OF SPECIFICATION\n\n")
  cat(sprintf("**%s demonstrates clinically significant %s interference on %s**\n\n",
              INTERFERENT, direction, ANALYTE))
  cat(sprintf("- Threshold concentration: ≥%.0f\n", threshold))
  cat(sprintf("- Maximum observed bias: %.2f%%\n", max_bias))
  cat(sprintf("- Acceptance criterion exceeded: ±%.1f%%\n\n", SPEC_LIMIT))
  cat("**Disposition:** Method validation FAILED for this interferent range.\n\n")
  cat("**Required Actions:**\n")
  cat("1. Document deviation per SOP-QA-001\n")
  cat("2. Perform root cause investigation\n")
  cat(sprintf("3. Implement sample rejection criteria for specimens with %s ≥%.0f\n",
              INTERFERENT, threshold))
  cat("4. Consider alternative analytical methodology\n")
} else if (nrow(borderline_rows) > 0) {
  cat("## RESULT: BORDERLINE - REVIEW REQUIRED\n\n")
  cat(sprintf("**%s shows borderline effects on %s measurement**\n\n", INTERFERENT, ANALYTE))
  cat("While the acceptance criterion was not exceeded, results warrant additional review.\n\n")
  cat("**Disposition:** Method validation CONDITIONAL PASS pending review.\n")
} else {
  cat("## RESULT: PASS\n\n")
  cat(sprintf("**No significant interference detected for %s on %s**\n\n", INTERFERENT, ANALYTE))
  cat(sprintf("All tested concentrations remain within the acceptance criterion of ±%.1f%%.\n\n", SPEC_LIMIT))
  cat("**Disposition:** Method validation PASSED for this interferent.\n")
}
```

# Study Design

## Objective

To evaluate the specificity of the `r ANALYTE` assay in the presence of `r INTERFERENT` per ICH Q2(R1) Section 2.2 (Specificity) requirements.

## Methodology

The study followed a dose-response design consistent with CLSI EP07-A3:

- **Interferent concentrations tested:** `r paste(sort(unique(filtered$interferent_concentration)), collapse = ", ")` `r filtered$interferent_unit[1]`
- **Pool levels:** `r POOL_LEVEL`
- **Replicates per condition:** `r filtered |> group_by(interferent_concentration) |> summarise(n = n()) |> pull(n) |> unique() |> paste(collapse = ", ")`
- **Acceptance criterion:** ±`r SPEC_LIMIT`% bias from baseline

## Materials

```{r materials, echo=FALSE}
tibble(
  Item = c("Analyte", "Interferent", "Instrument", "Reagent Lot", "Study Date Range"),
  Details = c(
    ANALYTE,
    INTERFERENT,
    paste(unique(filtered$instrument_id), collapse = ", "),
    paste(unique(filtered$lot_number)[1:3], collapse = ", "),
    paste(range(filtered$study_date), collapse = " to ")
  )
) |>
  gt() |>
  tab_header(title = "Study Materials") |>
  cols_width(Item ~ px(150), Details ~ px(300))
```

# Results

## Descriptive Statistics (gtsummary)

```{r stats-table, echo=FALSE}
filtered |>
  mutate(
    conc_label = paste0(interferent_concentration, " ", interferent_unit)
  ) |>
  select(conc_label, measured_value, bias_percent) |>
  tbl_summary(
    by = conc_label,
    statistic = list(
      measured_value ~ "{mean} ± {sd}",
      bias_percent ~ "{mean} ± {sd}"
    ),
    label = list(
      measured_value ~ "Measured Value",
      bias_percent ~ "Bias (%)"
    ),
    digits = list(
      measured_value ~ 3,
      bias_percent ~ 2
    )
  ) |>
  add_n() |>
  modify_caption(sprintf("**%s Interference on %s - Summary Statistics**", INTERFERENT, ANALYTE)) |>
  as_gt() |>
  tab_options(table.font.size = px(11))
```

## Dose-Response Curve

```{r dose-response-plot, echo=FALSE, fig.width=10, fig.height=6}
baseline_mean <- stats_df |> filter(interferent_concentration == 0) |> pull(mean_measured)

ggplot() +
  # Individual points
  geom_point(data = filtered,
             aes(x = interferent_concentration, y = measured_value),
             alpha = 0.4, color = "#00A3E0", size = 2) +
  # Mean line
  geom_line(data = stats_df,
            aes(x = interferent_concentration, y = mean_measured),
            color = "#003087", linewidth = 1.2) +
  geom_point(data = stats_df,
             aes(x = interferent_concentration, y = mean_measured),
             color = "#003087", size = 4) +
  # 95% CI ribbon
  geom_ribbon(data = stats_df,
              aes(x = interferent_concentration,
                  ymin = mean_measured - 1.96 * sd_measured / sqrt(n),
                  ymax = mean_measured + 1.96 * sd_measured / sqrt(n)),
              alpha = 0.2, fill = "#003087") +
  # Reference lines
  geom_hline(yintercept = baseline_mean, linetype = "dashed", color = "#00843D") +
  geom_hline(yintercept = baseline_mean * (1 + SPEC_LIMIT/100),
             linetype = "dotted", color = "#C8102E") +
  geom_hline(yintercept = baseline_mean * (1 - SPEC_LIMIT/100),
             linetype = "dotted", color = "#C8102E") +
  annotate("text", x = max(filtered$interferent_concentration) * 0.9,
           y = baseline_mean * 1.01, label = "Baseline", color = "#00843D", size = 3) +
  annotate("text", x = max(filtered$interferent_concentration) * 0.9,
           y = baseline_mean * (1 + SPEC_LIMIT/100) * 1.01,
           label = paste0("+", SPEC_LIMIT, "%"), color = "#C8102E", size = 3) +
  labs(
    title = sprintf("Interference of %s on %s", INTERFERENT, ANALYTE),
    subtitle = "Mean ± 95% CI",
    x = sprintf("%s (%s)", INTERFERENT, filtered$interferent_unit[1]),
    y = sprintf("%s (%s)", ANALYTE, filtered$analyte_unit[1])
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```

## Percent Recovery Plot

```{r recovery-plot, echo=FALSE, fig.width=10, fig.height=5}
stats_df |>
  mutate(
    recovery_pct = (mean_measured / baseline_mean) * 100,
    color = case_when(
      assessment == "OOS" ~ "#C8102E",
      assessment == "Borderline" ~ "#FF6D00",
      TRUE ~ "#00843D"
    )
  ) |>
  ggplot(aes(x = factor(interferent_concentration), y = recovery_pct, fill = color)) +
  geom_col(color = "black", width = 0.7) +
  geom_hline(yintercept = 100, color = "black", linewidth = 1) +
  geom_hline(yintercept = 100 - SPEC_LIMIT, linetype = "dashed", color = "#C8102E") +
  geom_hline(yintercept = 100 + SPEC_LIMIT, linetype = "dashed", color = "#C8102E") +
  scale_fill_identity() +
  labs(
    title = "Percent Recovery at Each Interferent Concentration",
    subtitle = sprintf("Acceptance: %.0f%% - %.0f%%", 100 - SPEC_LIMIT, 100 + SPEC_LIMIT),
    x = sprintf("%s Concentration (%s)", INTERFERENT, filtered$interferent_unit[1]),
    y = "% Recovery"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```

## Bias Analysis

```{r bias-plot, echo=FALSE, fig.width=10, fig.height=5}
stats_df |>
  mutate(
    color = case_when(
      assessment == "OOS" ~ "#C8102E",
      assessment == "Borderline" ~ "#FF6D00",
      assessment == "Baseline" ~ "#F5F5F5",
      TRUE ~ "#00843D"
    )
  ) |>
  ggplot(aes(x = factor(interferent_concentration), y = mean_bias_pct, fill = color)) +
  geom_col(color = "black", width = 0.7) +
  geom_errorbar(aes(ymin = mean_bias_pct - sd_bias_pct,
                    ymax = mean_bias_pct + sd_bias_pct),
                width = 0.2) +
  geom_hline(yintercept = 0, color = "black") +
  geom_hline(yintercept = SPEC_LIMIT, linetype = "dashed", color = "#C8102E") +
  geom_hline(yintercept = -SPEC_LIMIT, linetype = "dashed", color = "#C8102E") +
  annotate("rect", xmin = -Inf, xmax = Inf,
           ymin = -SPEC_LIMIT, ymax = SPEC_LIMIT,
           fill = "#00843D", alpha = 0.1) +
  scale_fill_identity() +
  labs(
    title = "Percent Bias at Each Interferent Concentration",
    subtitle = sprintf("Acceptance limits: ±%.1f%%", SPEC_LIMIT),
    x = sprintf("%s Concentration (%s)", INTERFERENT, filtered$interferent_unit[1]),
    y = "Bias (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```

## Assessment Summary

```{r assessment-table, echo=FALSE}
stats_df |>
  select(interferent_concentration, n, mean_measured, sd_measured, cv_percent, mean_bias_pct, assessment) |>
  gt() |>
  cols_label(
    interferent_concentration = "Concentration",
    n = "N",
    mean_measured = "Mean",
    sd_measured = "SD",
    cv_percent = "CV%",
    mean_bias_pct = "Bias%",
    assessment = "Result"
  ) |>
  fmt_number(columns = c(mean_measured, sd_measured), decimals = 3) |>
  fmt_number(columns = c(cv_percent, mean_bias_pct), decimals = 2) |>
  data_color(
    columns = assessment,
    colors = scales::col_factor(
      palette = c("#F5F5F5", "#FF6D00", "#C8102E", "#00843D"),
      domain = c("Baseline", "Borderline", "OOS", "Pass")
    )
  ) |>
  tab_header(
    title = "Interference Assessment Results",
    subtitle = sprintf("Acceptance: ±%.1f%%", SPEC_LIMIT)
  )
```

# Statistical Analysis

## TOST Equivalence Test

Two One-Sided Tests (TOST) procedure to demonstrate statistical equivalence within ±`r EQUIV_MARGIN`% margin:

```{r tost-analysis, echo=FALSE}
baseline_data <- filtered |> filter(interferent_concentration == 0)
test_data <- filtered |> filter(interferent_concentration > 0)

if (nrow(baseline_data) >= 3 && nrow(test_data) >= 3) {
  tost_results <- test_data |>
    group_by(interferent_concentration) |>
    summarise(
      n_test = n(),
      mean_test = mean(measured_value),
      sd_test = sd(measured_value),
      .groups = "drop"
    ) |>
    mutate(
      baseline_mean = mean(baseline_data$measured_value),
      baseline_sd = sd(baseline_data$measured_value),
      n_baseline = nrow(baseline_data),
      diff = mean_test - baseline_mean,
      diff_pct = (diff / baseline_mean) * 100,
      pooled_se = sqrt(sd_test^2 / n_test + baseline_sd^2 / n_baseline),
      t_lower = (diff - (-EQUIV_MARGIN / 100 * baseline_mean)) / pooled_se,
      t_upper = (diff - (EQUIV_MARGIN / 100 * baseline_mean)) / pooled_se,
      df = n_test + n_baseline - 2,
      p_lower = 1 - pt(t_lower, df),
      p_upper = pt(t_upper, df),
      equivalent = p_lower < 0.05 & p_upper < 0.05
    )

  tost_results |>
    select(interferent_concentration, n_test, diff_pct, p_lower, p_upper, equivalent) |>
    gt() |>
    cols_label(
      interferent_concentration = "Concentration",
      n_test = "N",
      diff_pct = "Diff (%)",
      p_lower = "P (lower)",
      p_upper = "P (upper)",
      equivalent = "Equivalent"
    ) |>
    fmt_number(columns = c(diff_pct), decimals = 2) |>
    fmt_number(columns = c(p_lower, p_upper), decimals = 4) |>
    data_color(
      columns = equivalent,
      colors = scales::col_factor(
        palette = c("#C8102E", "#00843D"),
        domain = c(FALSE, TRUE)
      )
    ) |>
    tab_header(
      title = "TOST Equivalence Results",
      subtitle = sprintf("α = 0.05, Equivalence margin = ±%.1f%%", EQUIV_MARGIN)
    )
} else {
  cat("Insufficient data for TOST analysis.")
}
```

## Linear Regression

```{r regression, echo=FALSE}
non_baseline <- stats_df |> filter(interferent_concentration > 0)

if (nrow(non_baseline) >= 3) {
  model <- lm(mean_bias_pct ~ interferent_concentration, data = non_baseline)

  tidy(model) |>
    gt() |>
    fmt_number(columns = c(estimate, std.error, statistic), decimals = 4) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title = "Linear Regression: Bias vs Concentration",
      subtitle = sprintf("R² = %.4f", summary(model)$r.squared)
    )
}
```

# Conclusion

```{r conclusion, echo=FALSE, results='asis'}
pass_count <- sum(stats_df$assessment == "Pass")
borderline_count <- sum(stats_df$assessment == "Borderline")
oos_count <- sum(stats_df$assessment == "OOS")
total_tested <- pass_count + borderline_count + oos_count

cat("## Study Disposition\n\n")
cat(sprintf("| Metric | Value |\n"))
cat("|--------|-------|\n")
cat(sprintf("| Concentrations Tested | %d |\n", total_tested + 1))
cat(sprintf("| Pass | %d |\n", pass_count))
cat(sprintf("| Borderline | %d |\n", borderline_count))
cat(sprintf("| OOS | %d |\n", oos_count))
cat(sprintf("| Acceptance Criterion | ±%.1f%% |\n\n", SPEC_LIMIT))

if (oos_count > 0) {
  cat("**OVERALL RESULT: FAIL**\n\n")
  cat("The analytical method does not meet ICH Q2(R1) specificity requirements for the tested interferent range.\n")
} else {
  cat("**OVERALL RESULT: PASS**\n\n")
  cat("The analytical method meets ICH Q2(R1) specificity requirements. ")
  cat(sprintf("%s does not significantly interfere with %s quantitation across the tested concentration range.\n",
              INTERFERENT, ANALYTE))
}
```

# Signatures

| Role | Name | Signature | Date |
|------|------|-----------|------|
| **Performed By** | _________________ | _________________ | ________ |
| **Reviewed By** | _________________ | _________________ | ________ |
| **Approved By** | _________________ | _________________ | ________ |

---

# Appendix A: Raw Data (First 20 Records)

```{r raw-data, echo=FALSE}
filtered |>
  head(20) |>
  select(interferent_concentration, pool_level, replicate, baseline_value,
         measured_value, bias_percent, technician_id, instrument_id) |>
  gt() |>
  fmt_number(columns = c(baseline_value, measured_value), decimals = 3) |>
  fmt_number(columns = bias_percent, decimals = 2) |>
  tab_header(title = "Raw Study Data (Sample)")
```

---

*Report generated automatically using Posit Quarto per protocol `r PROTOCOL`.*

*Analysis conducted per ICH Q2(R1) and CLSI EP07-A3 guidelines.*

*This document is electronically generated. Signatures above constitute approval.*
